Sexual selection improves population fitness: a meta-analysis

### 2019-01-02 Response_to_Nature_Comms.Rmd

Many thanks for your time and for you very helpful feedback.

>*Firstly, it seems that several male traits like to be under sexual selection, but not related to population fitness were included. I don’t think this is correct given the question being asked (i.e. whether sexual selection increases population fitness). I think traits which are directly under sexual selection (eg. attractiveness, reproductive success) should not be pooled with traits which may respond to manipulation of sexual selection indirectly (eg. male development time, survival) and can affect population fitness. Distinguishing between both types of trait could actually be revealing – e.g. could expose trade-offs between sexually selected traits and fitness components unrelated to reproductive competition (see eg. Radwan et al. 2015 Evol Biol), a thus potentially explain lower effect of sexual selection on (pooled) male traits than on female traits.*

We completely agree that male attractiveness does not affect population fitness directly, and it is important to separate this trait from the "true" measures of population fitness in the meta-analysis. We do this more carefully in the revised meta-analysis. Specifically, in the original study, we classified male attractiveness as having an "indirect" relationship with population fitness (this is also how we classified traits such as body condition and life span; see Table S1 for a full breakdown). However on reflection, we think that attractiveness is only tenuously linked to population fitness, and so we have re-classified its relationship with population fitness as "ambiguous", along with traits like "Immunity" and "Development rate". Our main analysis excludes all of the ambiguous traits, and so our main results are now not derived from any effect sizes relating to male attractiveness. The exclusion of the n = 6 attractiveness effect sizes did not qualitatively change the results.

Additionally, we have split the trait "Reproductive success" into "Female Reproductive Success", "Male Reproductive Success" and "Both Reproductive Success", the latter from studies where offspring number was measured from a given male-female pair interaction. We have made this reclassification in recognition of the facts that these two fitness components are maximised very differently, and are likely to have differentially related to population fitness. Specifically, we maintain "Female Reproductive Success" and "Both Reproductive Success" as direct measures of fitness but class "Male Reproductive Success" as an ambiguous measure of fitness.

### 2019-01-02 meta_analysis_sexual_selection_pop_fitness.Rmd
Updated Reproductive Success

```{r, fig.height=18, fig.width=12}
# Create new factor to order factors in a way where Ambig, Indirect and Direct are Grouped
full_dataset$Outcome_f = factor(full_dataset$Outcome, levels = c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Extinction Rate', 'Offspring Viability', 'Reproductive Success'))

@@

knitr::include_graphics(path = "figures/ForestPlot_size_reduced.png") 
```

@@

outcome.list.factor <-  c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Extinction Rate', 'Offspring Viability', 'Reproductive Success')

@@

# Obtain Bayes Factor of likelihood that the outcome is greater than 0
BF.outcome.list <-  c('Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Extinction Rate', 'Offspring Viability', 'Reproductive Success')
# Obtain BF for all components
Hypotheses <- list()
BFs <- list()
Hypotheses[["Behavioural Plasticity"]] <- hypothesis(components.brms, "Intercept > 0") #Intercept in model, need to do it by itself
BFs[["Behavioural Plasticity"]] <- Hypotheses[["Behavioural Plasticity"]][["hypothesis"]][["Evid.Ratio"]]
for (i in BF.outcome.list){
  Hypotheses[[i]] <- hypothesis(components.brms, paste("Intercept + Outcome", i," > 0", sep = ""))
  BFs[[i]] <- Hypotheses[[i]][["hypothesis"]][["Evid.Ratio"]]
} #Loop over all fitness compoinents

@@

get.predictions.outcomes <- function(newdata){
  `Body Condition`<-0; `Body Size`<-0; `Development Rate`<-0; `Early Fecundity`<-0; `Ejaculate Quality and Production`<-0; `Extinction Rate`<-0; `Fitness Senescence`<-0; Immunity<-0; Lifespan<-0; `Male Attractiveness`<-0; `Mating Duration`<-0; `Mating Frequency`<-0; `Mating Latency`<-0; `Mating Success`<-0; `Mutant Frequency`<-0; `Offspring Viability`<-0; `Pesticide Resistance`<-0; `Reproductive Success`<-0; Strength<-0
  if(newdata[1] == 'Body Size') `Body Size`<-1
  if(newdata[1] == 'Development Rate') `Development Rate`<-1
  if(newdata[1] == 'Early Fecundity') `Early Fecundity`<-1
  if(newdata[1] == 'Immunity')Immunity<-1
  if(newdata[1] == 'Mating Duration')`Mating Duration`<-1
  if(newdata[1] == 'Pesticide Resistance')`Pesticide Resistance`<-1
  if(newdata[1] == 'Mutant Frequency')`Mutant Frequency`<-1
  if(newdata[1] == 'Body Condition')`Body Condition`<-1
  if(newdata[1] == 'Fitness Senescence')`Fitness Senescence`<-1
  if(newdata[1] == 'Lifespan')Lifespan<-1
  if(newdata[1] == 'Male Attractiveness')`Male Attractiveness`<-1
  if(newdata[1] == 'Mating Frequency')`Mating Frequency`<-1
  if(newdata[1] == 'Mating Latency')`Mating Latency`<-1
  if(newdata[1] == 'Mating Success')`Mating Success`<-1
  if(newdata[1] == 'Strength')Strength<-1
  if(newdata[1] == 'Ejaculate Quality and Production')`Ejaculate Quality and Production`<-1
  if(newdata[1] == 'Extinction Rate')`Extinction Rate`<-1
  if(newdata[1] == 'Offspring Viability')`Offspring Viability`<-1
  if(newdata[1] == 'Reproductive Success')`Reproductive Success`<-1

  predict(model.fitness.components.REML, newmods=c(`Body Condition`, `Body Size`, `Development Rate`, `Early Fecundity`, `Ejaculate Quality and Production`, `Extinction Rate`, `Fitness Senescence`, Immunity, Lifespan, `Male Attractiveness`, `Mating Duration`, `Mating Frequency`, `Mating Latency`, `Mating Success`, `Mutant Frequency`, `Offspring Viability`, `Pesticide Resistance`, `Reproductive Success`, Strength))
}
# Get the predictions for each combination of moderators
predictions.outcomes <- as.data.frame(expand.grid(Outcome = outcome.list.factor))
predictions.outcomes <- cbind(predictions.outcomes, do.call("rbind", apply(predictions.outcomes, 1, get.predictions.outcomes))) %>%
  select(Outcome, pred, se, ci.lb, ci.ub) 
for(i in 2:5) predictions.outcomes[,i] <- unlist(predictions.outcomes[,i])

@@

```{r, warning = FALSE}
outcome.list <- as.list(c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Extinction Rate', 'Offspring Viability', 'Reproductive Success'))
names(outcome.list) <- c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Extinction Rate', 'Offspring Viability', 'Reproductive Success')
outcome.models <-  llply(outcome.list, function(x) rma.mv(g, var.g, 
                          mods = ~ 1, 
                          method = "REML",
                          random = list(~ 1 | Study.ID,
                                       ~ 1 | Taxon),
                          subset = (Outcome_f == x),
                          data = full_dataset))

@@

@@

#Reproductive Success
restricted.dataRS <- full_dataset %>% filter(full_dataset$Outcome == "Reproductive Success")

#Run estimate of heterogeneity
W = diag(1/restricted.dataRS$var.g)
X = model.matrix(outcome.models[["Reproductive Success"]])
P = W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
ReproductiveSuccessI2 <- 100 * sum(outcome.models[["Reproductive Success"]]$sigma2) / (sum(outcome.models[["Reproductive Success"]]$sigma2) + (outcome.models[["Reproductive Success"]]$k-outcome.models[["Reproductive Success"]]$p)/sum(diag(P)))

simple.frame$I2 <- c(NA, BodySizeI2, DevelopmentRateI2, EarlyFecundityI2, ImmunityI2, MaleAttractivenessI2, MatingDurationI2, NA, MutantFrequencyI2, FitnessSenescenceI2, LifespanI2, MatingFrequencyI2, MatingLatencyI2, MatingSuccessI2, NA, EjaculateQualityI2, ExtinctionRateI2, OffspringViabilityI2, ReproductiveSuccessI2)

outcome.frame <- format(simple.frame, digits = 2)
outcome.frame <- add_rownames(outcome.frame, "Outcome")
outcome.frame$b <- as.numeric(outcome.frame$b)
outcome.frame$k <- as.numeric(outcome.frame$k)
outcome.frame$ci.lb <- as.numeric(outcome.frame$ci.lb)
outcome.frame$ci.ub <- as.numeric(outcome.frame$ci.ub)
outcome.frame$I2 <- as.numeric(outcome.frame$I2)
outcome.frame$Class <- c("Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Direct", "Direct", "Direct")

@@

```{r}
model.preliminary <- rma.mv(g, var.g, 
                         mods = ~ 1 + Sex * Environment + Taxon + Outcome.Class + log(Generations) + Blinding, 
                         random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome), 
                         method = "REML", 
                         data = full_dataset)

@@

Here we can also run a Bayesian model alongside the REML model (metafor). The _R^2^_ for this model is 0.36 (95% CIs = 0.33-0.4). 
```{r, warning = FALSE, results='hide', message=FALSE, error = FALSE}
if(!file.exists("data/brms.preliminary.rds")){
  brms.preliminary <- brm(g | se(SE)  ~ 1 + Sex * Environment + log(Generations) + Blinding #Note that running se(SE, sigma = TRUE) gives different result due to a difference in priors
                + (1|Study.ID) #group level effects
                + (1|Outcome)
                + (1|Taxon), 
                family = "gaussian", 
                seed = 1,
                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
                control = list(adapt_delta = 0.999, max_treedepth = 15),
                data = full_dataset %>% mutate(SE = sqrt(var.g)))
  saveRDS(brms.preliminary, file = "data/brms.preliminary.rds") 
}
brms.preliminary <- readRDS(file = "data/brms.preliminary.rds") # Avoid re-running model above
```
<br></br>

>>>>

```{r, fig.height=18, fig.width=12}
# Create new factor to order factors in a way where Ambig, Indirect and Direct are Grouped
full_dataset$Outcome_f = factor(full_dataset$Outcome, levels = c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Male Reproductive Success', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Both Reproductive Success', 'Extinction Rate', 'Female Reproductive Success', 'Offspring Viability'))

@@

#knitr::include_graphics(path = "figures/ForestPlot_size_reduced.png") 
```

@@

outcome.list.factor <-  c('Behavioural Plasticity', 'Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Male Reproductive Success', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Both Reproductive Success', 'Extinction Rate', 'Female Reproductive Success', 'Offspring Viability')

@@

# Obtain Bayes Factor of likelihood that the outcome is greater than 0
BF.outcome.list <-  c('Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Male Reproductive Success', 'Mating Duration', 'Pesticide Resistance', 'Mutant Frequency', 'Body Condition', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Strength', 'Ejaculate Quality and Production', 'Both Reproductive Success', 'Extinction Rate', 'Female Reproductive Success', 'Offspring Viability')
# Obtain BF for all components
Hypotheses <- list()
BFs <- list()
Hypotheses[["Behavioural Plasticity"]] <- hypothesis(components.brms, "Intercept > 0") #Intercept in model, need to do it by itself
BFs[["Behavioural Plasticity"]] <- Hypotheses[["Behavioural Plasticity"]][["hypothesis"]][["Evid.Ratio"]]
for (i in BF.outcome.list){
  Hypotheses[[i]] <- hypothesis(components.brms, paste("Intercept + Outcome", i," > 0", sep = ""))
  BFs[[i]] <- Hypotheses[[i]][["hypothesis"]][["Evid.Ratio"]]
} #Loop over all fitness compoinents

@@

get.predictions.outcomes <- function(newdata){
  `Body Condition`<-0; `Body Size`<-0; `Development Rate`<-0; `Early Fecundity`<-0; `Ejaculate Quality and Production`<-0; `Extinction Rate`<-0; `Fitness Senescence`<-0; Immunity<-0; Lifespan<-0; `Male Attractiveness`<-0; `Male Reproductive Success`<-0; `Mating Duration`<-0; `Mating Frequency`<-0; `Mating Latency`<-0; `Mating Success`<-0; `Mutant Frequency`<-0; `Offspring Viability`<-0; `Pesticide Resistance`<-0; `Female Reproductive Success`<-0; `Both Reproductive Success`<-0; Strength<-0
  if(newdata[1] == 'Body Size') `Body Size`<-1
  if(newdata[1] == 'Development Rate') `Development Rate`<-1
  if(newdata[1] == 'Early Fecundity') `Early Fecundity`<-1
  if(newdata[1] == 'Immunity')Immunity<-1
  if(newdata[1] == 'Mating Duration')`Mating Duration`<-1
  if(newdata[1] == 'Pesticide Resistance')`Pesticide Resistance`<-1
  if(newdata[1] == 'Mutant Frequency')`Mutant Frequency`<-1
  if(newdata[1] == 'Body Condition')`Body Condition`<-1
  if(newdata[1] == 'Fitness Senescence')`Fitness Senescence`<-1
  if(newdata[1] == 'Lifespan')Lifespan<-1
  if(newdata[1] == 'Male Attractiveness')`Male Attractiveness`<-1
  if(newdata[1] == 'Male Reproductive Success')`Male Reproductive Success`<-1
  if(newdata[1] == 'Mating Frequency')`Mating Frequency`<-1
  if(newdata[1] == 'Mating Latency')`Mating Latency`<-1
  if(newdata[1] == 'Mating Success')`Mating Success`<-1
  if(newdata[1] == 'Strength')Strength<-1
  if(newdata[1] == 'Ejaculate Quality and Production')`Ejaculate Quality and Production`<-1
  if(newdata[1] == 'Extinction Rate')`Extinction Rate`<-1
  if(newdata[1] == 'Offspring Viability')`Offspring Viability`<-1
  if(newdata[1] == 'Female Reproductive Success')`Female Reproductive Success`<-1
  if(newdata[1] == 'Both Reproductive Success')`Both Reproductive Success`<-1

  predict(model.fitness.components.REML, newmods=c(`Body Condition`, `Body Size`, `Development Rate`, `Early Fecundity`, `Ejaculate Quality and Production`, `Extinction Rate`, `Fitness Senescence`, Immunity, Lifespan, `Male Attractiveness`, `Male Reproductive Success`, `Mating Duration`, `Mating Frequency`, `Mating Latency`, `Mating Success`, `Mutant Frequency`, `Offspring Viability`, `Pesticide Resistance`, `Female Reproductive Success`, `Both Reproductive Success`, Strength))
}
# Get the predictions for each combination of moderators
predictions.outcomes <- as.data.frame(expand.grid(Outcome = outcome.list.factor))
predictions.outcomes <- cbind(predictions.outcomes, do.call("rbind", apply(predictions.outcomes, 1, get.predictions.outcomes))) %>%
  select(Outcome, pred, se, ci.lb, ci.ub) 
for(i in 2:5) predictions.outcomes[,i] <- unlist(predictions.outcomes[,i])

@@

```{r, warning = FALSE}
#Excluding those with 3 or less effect sizes.

outcome.list <- as.list(c('Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Male Reproductive Success', 'Mating Duration', 'Mutant Frequency', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Ejaculate Quality and Production', 'Both Reproductive Success', 'Extinction Rate', 'Female Reproductive Success', 'Offspring Viability'))
names(outcome.list) <- c('Body Size', 'Development Rate', 'Early Fecundity', 'Immunity', 'Male Attractiveness', 'Male Reproductive Success', 'Mating Duration', 'Mutant Frequency', 'Fitness Senescence', 'Lifespan', 'Mating Frequency', 'Mating Latency', 'Mating Success', 'Ejaculate Quality and Production', 'Both Reproductive Success', 'Extinction Rate', 'Female Reproductive Success', 'Offspring Viability')
outcome.models <-  llply(outcome.list, function(x) rma.mv(g, var.g, 
                          mods = ~ 1, 
                          method = "REML",
                          random = list(~ 1 | Study.ID,
                                       ~ 1 | Taxon),
                          subset = (Outcome_f == x),
                          data = full_dataset))

@@

#Male Reproductive Success
restricted.dataMRS <- full_dataset %>% filter(full_dataset$Outcome == "Male Reproductive Success")

#Run estimate of heterogeneity
W = diag(1/restricted.dataMRS$var.g)
X = model.matrix(outcome.models[["Male Reproductive Success"]])
P = W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
MaleReproductiveSuccessI2 <- 100 * sum(outcome.models[["Male Reproductive Success"]]$sigma2) / (sum(outcome.models[["Male Reproductive Success"]]$sigma2) + (outcome.models[["Male Reproductive Success"]]$k-outcome.models[["Male Reproductive Success"]]$p)/sum(diag(P)))

@@

#Both Reproductive Success
restricted.dataBRS <- full_dataset %>% filter(full_dataset$Outcome == "Both Reproductive Success")

#Run estimate of heterogeneity
W = diag(1/restricted.dataBRS$var.g)
X = model.matrix(outcome.models[["Both Reproductive Success"]])
P = W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
BothReproductiveSuccessI2 <- 100 * sum(outcome.models[["Both Reproductive Success"]]$sigma2) / (sum(outcome.models[["Both Reproductive Success"]]$sigma2) + (outcome.models[["Both Reproductive Success"]]$k-outcome.models[["Both Reproductive Success"]]$p)/sum(diag(P)))

#Female Reproductive Success
restricted.dataFRS <- full_dataset %>% filter(full_dataset$Outcome == "Female Reproductive Success")

#Run estimate of heterogeneity
W = diag(1/restricted.dataFRS$var.g)
X = model.matrix(outcome.models[["Female Reproductive Success"]])
P = W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
FemaleReproductiveSuccessI2 <- 100 * sum(outcome.models[["Female Reproductive Success"]]$sigma2) / (sum(outcome.models[["Female Reproductive Success"]]$sigma2) + (outcome.models[["Female Reproductive Success"]]$k-outcome.models[["Female Reproductive Success"]]$p)/sum(diag(P)))

simple.frame$I2 <- c(BodySizeI2, DevelopmentRateI2, EarlyFecundityI2, ImmunityI2, MaleAttractivenessI2, MaleReproductiveSuccessI2, MatingDurationI2, MutantFrequencyI2, FitnessSenescenceI2, LifespanI2, MatingFrequencyI2, MatingLatencyI2, MatingSuccessI2, EjaculateQualityI2, BothReproductiveSuccessI2, ExtinctionRateI2, FemaleReproductiveSuccessI2, OffspringViabilityI2)

outcome.frame <- format(simple.frame, digits = 2)
outcome.frame <- add_rownames(outcome.frame, "Outcome")
outcome.frame$b <- as.numeric(outcome.frame$b)
outcome.frame$k <- as.numeric(outcome.frame$k)
outcome.frame$ci.lb <- as.numeric(outcome.frame$ci.lb)
outcome.frame$ci.ub <- as.numeric(outcome.frame$ci.ub)
outcome.frame$I2 <- as.numeric(outcome.frame$I2)
outcome.frame$Class <- c("Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Ambiguous", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Indirect", "Direct", "Direct", "Direct", "Direct")

@@

```{r}
model.preliminary <- rma.mv(g, var.g, 
                         mods = ~ 1 + Sex * Environment + Taxon + Outcome.Class + log(Generations) + Blinding + Enforced.Monogamy, 
                         random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome), 
                         method = "REML", 
                         data = full_dataset)

@@

Here we can also run a Bayesian model alongside the REML model (metafor). The _R^2^_ for this model is 0.36 (95% CIs = 0.33-0.4). 
```{r, warning = FALSE, results='hide', message=FALSE, error = FALSE}
if(!file.exists("data/brms.preliminary.rds")){
  brms.preliminary <- brm(g | se(SE)  ~ 1 + Sex * Environment + log(Generations) + Blinding + Enforced.Monogamy #Note that running se(SE, sigma = TRUE) gives different result due to a difference in priors
                + (1|Study.ID) #group level effects
                + (1|Outcome)
                + (1|Taxon), 
                family = "gaussian", 
                seed = 1,
                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
                control = list(adapt_delta = 0.999, max_treedepth = 15),
                data = full_dataset %>% mutate(SE = sqrt(var.g)))
  saveRDS(brms.preliminary, file = "data/brms.preliminary.rds") 
}
brms.preliminary <- readRDS(file = "data/brms.preliminary.rds") # Avoid re-running model above
```
<br></br>
